Extracellular potentials generated by axonal projections are shaped by patterns of bifurcations and terminations

### 2016-03-15 doc/manuscript.tex
add toods

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Extracellular potentials generated by axonal projections are shaped by patterns of bifurcations and terminations},
            pdfauthor={Thomas McColgan, Hermann Wagner, Catherine E Carr, Richard Kempter},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{setspace}
\doublespacing

\title{Extracellular potentials generated by axonal projections are shaped by
patterns of bifurcations and terminations}
\author{Thomas McColgan, Hermann Wagner, Catherine E Carr, Richard Kempter}
\date{\today}

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_1.pdf}
\caption{Axonal projections generate a dipole-like extracellular field
potential (EFP) when stimulated with a pulse of activity. \textbf{A}
shows the modeled axon bundle in black, along with iso-amplitude lines
for the low-pass filtered EFP signature of the activity pulse. The
contours show the typical double-lobe of a dipole. The waveforms,
recorded at the locations of the colored dots in \textbf{A} and shown in
\textbf{B}, also show a polarity reversal. The most salient locations
are highlighted with the red box. The transition ocurrs by inverting the
envelope with aproximately constant phase, not by a shifting of phase
with constant envelope. \textbf{C} shows the progression of the maximum
amplitude with depth (red line). It closely follows the local density of
bifurcations and terminations (black histogram). The situation is
different for the high-frequency component. The iso-amlitude contours
shown in \textbf{D} do not exhibit the same double-lobe structure, and
the waveforms (\textbf{C}) also don't reverse polarity. The amplitude
envelope of the high-frequency component (red line in \textbf{E})
follows the density of fibers (black histogram in \textbf{E}).}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_2.pdf}
\caption{Low-frequency component of the axon bundle EFP exceeds
high-frequency component in reach, modulated by axon structure. All
plots show the scaling of the low-frequency (blue lines) and
high-frequency (green lines) frequency components on a
doubly-logarithmic scale. Amplitudes were normalized to unit value at
the end of the tuft in axial direction (\textbf{A} and \textbf{B}) and
from the trunk in radial direction (\textbf{C} and \textbf{D}).
\textbf{A} and \textbf{C} show the behaviour in the case of a
bifurcating bundle, while \textbf{B} and \textbf{D} show the
nonbifurcating case. In the bifurcating case, the low-frequency
component exceeds the reach of the high-frequency comonent in both
directions. In the non-bifurcating case, the decay is the same for both
components in the axial direction. This is consistent with the presence
of a low-frequency dipole-like potential in the bifurcating bundle but
not in the non-bifurcating bundle.}
\end{figure}

\emph{The barn owl neurophonic as an example that shows these
properties.} In order to test our model of the extracellular field of
axon bundles, we compared it to recordings from the barn owl auditory
brain stem. We adjusted the model parameters to match the anatomy and
physiology of nucleus laminaris (NL), and set the activation statistics
to match measured PSTHs in response to stimulation by a click sound
presented to one ear. We then simulated the resulting EFP at varying
depths, corresponding to recording depths in the experiment. The
resulting recordings from experiment and simulation are shown in Figure
3.

@@

\begin{itemize}
\tightlist
\item
  Add simplified explanations of:
\item
  spatial filtering
\item
  dipole/stationary potentials
\end{itemize}

@@

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Extracellular potentials generated by axonal projections are shaped by patterns of bifurcations and terminations},
            pdfauthor={Thomas McColgan, Hermann Wagner, Ji Liu, Paula T Kuokkanen, Catherine E Carr, Richard Kempter},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{setspace}
\doublespacing

\title{Extracellular potentials generated by axonal projections are shaped by
patterns of bifurcations and terminations}
\author{Thomas McColgan, Hermann Wagner, Ji Liu, Paula T Kuokkanen, Catherine E
Carr, Richard Kempter}
\date{\today}

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_1.pdf}
\caption{Axonal projections generate a dipole-like extracellular field
potential (EFP) when stimulated with a pulse of activity. \textbf{A}
Modeled axon bundle in black, along with iso-amplitude lines for the
low-pass filtered EFP signature of the activity pulse. The contours
(maximum amplitude in mV as indicated by colorbar) show the typical
double-lobe of a dipole. \textbf{B} The waveforms, recorded at the
locations of the colored dots in \emph{A} also show a polarity reversal
(red box). The reversal ocurrs by inverting the envelope with
aproximately constant phase, not by a shifting of phase with constant
envelope. \textbf{C} Progression of the maximum amplitude with depth
(red line). It closely follows the local difference in density between
bifurcations and terminations (black histogram). \textbf{D} Iso-amlitude
contours for the high-frequency component, which do not exhibit a
doiuble-lobed structure. \textbf{E} Response waveforms of the
low-frequency component. The responses do not reverse their polarity.
\textbf{F} Maximum amplitude of the high-frequency component (red line)
follows the density of fibers (black histogram).}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_2.pdf}
\caption{Low-frequency component of the axon bundle EFP exceeds
high-frequency component in reach, modulated by axon structure. All
plots show the scaling of the low-frequency (blue lines) and
high-frequency (green lines) frequency components on a
doubly-logarithmic scale. Amplitudes were normalized to unit value at
the end of the tuft in axial direction (\textbf{A} and \textbf{B}) and
from the trunk in radial direction (\textbf{C} and \textbf{D}).
\textbf{A} and \textbf{C} show the behaviour in the case of a
bifurcating bundle, while \textbf{B} and \textbf{D} show the
nonbifurcating case. In the bifurcating case, the low-frequency
component exceeds the reach of the high-frequency comonent in both
directions. In the non-bifurcating case, the decay is the same for both
components in the axial direction. The scaling behaviours are consistent
with the presence of a low-frequency dipole-like potential in the
bifurcating bundle but not in the non-bifurcating bundle.}
\end{figure}

\emph{The barn owl neurophonic as an example that shows these
properties.} In order to test our model of the extracellular field of
axon bundles, we compared it to recordings from the barn owl auditory
brain stem.

We adjusted the model parameters to match the anatomy and physiology of
nucleus laminaris (NL), and set the activation statistics to match
measured PSTHs in response to stimulation by a click sound presented to
one ear. We then simulated the resulting EFP at varying depths,
corresponding to recording depths in the experiment. The resulting
recordings from experiment and simulation are shown in Figure 3.

@@

@@

\section{Todos}\label{todos}

\begin{itemize}
\tightlist
\item
  Code

  \begin{itemize}
  \tightlist
  \item
    implement new simplification
  \item
    adapt generic bundle params to be more typical
  \item
    make click psth more generic
  \end{itemize}
\item
  Introduction

  \begin{itemize}
  \tightlist
  \item
    find more examples in literature
  \end{itemize}
\item
  Results

  \begin{itemize}
  \tightlist
  \item
    Page 4 changes
  \item
    much more substantial explanation of Fig 2
  \item
    describe differences between barn owl and generic model
  \item
    Add more biological motivation for figure 2 normalization
  \end{itemize}
\item
  Figures

  \begin{itemize}
  \tightlist
  \item
    New Figure 1

    \begin{itemize}
    \tightlist
    \item
      single axons

      \begin{itemize}
      \tightlist
      \item
        straight
      \item
        branching
      \item
        terminating
      \end{itemize}
    \item
      compound
    \item
      CSD
    \item
      PSTH
    \end{itemize}
  \item
    Figure 1

    \begin{itemize}
    \tightlist
    \item
      version with MUA
    \item
      lower density of axons
    \item
      Add scalebar for time
    \item
      change color of axis label
    \item
      change color of box
    \item
      add PSTH
    \item
      align zeros in C
    \item
      add contour labels?
    \end{itemize}
  \item
    Figure 2

    \begin{itemize}
    \tightlist
    \item
      add diagrams
    \item
      space between B and C
    \item
      better location for axial scaling
    \end{itemize}
  \item
    Figure 5Bb log scale
  \end{itemize}
\item
  Methods

  \begin{itemize}
  \tightlist
  \item
    Adapt analytical model to new method
  \item
    Add simplified explanations of:

    \begin{itemize}
    \tightlist
    \item
      spatial filtering
    \item
      dipole/stationary potentials
    \end{itemize}
  \item
    relationship to CSD
  \end{itemize}
\item
  Discussion
\item
  Abstract
\end{itemize}
